Effects of rotation on the Lyman-$\alpha$ line morphology in distant galaxies

### 2013-02-27 paper/Analisis.tex
heavy editting

\begin{table}
\begin{center}
\begin{tabular}{ccc}\hline
Velocity (\kms) & $V_{\rm max}$&$0,\ 50,\ 100,\ 200,\ 300$\\
Hydrogen Optical Depth & $\tau_{H} $ & $10^{5},\ 10^{6},\ 10^{7}$\\
Dust Optical Depth & $\tau_{A}$ & XXX \\
Photons Distributions & & Central, Homogeneous\\
\hline
\end{tabular}
\caption{
Values for the varying input parameters in CLARA. Taking into account
all the possible combinations for these models
} 
\label{table:models}
\end{center}
\end{table}

@@

The central result of this paper is summarized in Figure
~\ref{fig:HOMandCentral} where we show that rotation as a considerable
effect on the morphology of the emergent Lyman-alpha line both in the
case where the photons are emitted at the sphere's center and when
they are initialized with an homogeneous distribution all over the
gas volume.

The results for this outgoing spectra are integrated over the whole
sphere, meaning that all the escaping photons were taking into account
regardless of the direction of the outgoing photons. Figure
\ref{fig:DifferentObservers} shows how if one gives a weight to each outgoing
photon according to its direction when escaping the gas distribution
it is posible to detect notable differences in the spectrum for
different viewing angles.

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{7tDifSpeedsZ.png}
  \includegraphics[width=0.45\textwidth]{7tHOMDifSpeeds1.png}
  \label{fig:HOMandCentral}\caption{Shape of the lyman alpha line for
    differents velocities. The left (right) panel shows the central
    (homogeneous) photon distribution.}
\end{figure*}

\begin{figure*}
\includegraphics[scale=0.45]{Observers7t.png}
\includegraphics[scale=0.45]{Observers7tHOM.png}
\caption{Spectra for different observers. Model:$V=300$\kms, Optical
  Depth $\tau=10^{7}$ and Central Distribution without dust.} 
  \label{fig:DifferentObservers}
\end{figure*}

\subsection{Maximums Position}
\label{sec:MP}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{maximumvsODDifSpeedsCentral.png}
    \includegraphics[width=0.45\textwidth]{maximumvsODDifSpeedsHOM.png}
  \label{figure:maximavsOD}\caption{Position of the maxima in the outgoing spectra for differents Optical Depths, (up)Central Distriution, (Down) Homogeneous Distribution.}
\end{figure}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{maximumvsvelocitiesDifODCentral.png}
    \includegraphics[width=0.45\textwidth]{maximumvsvelocitiesDifODHOM.png}
  \label{figure:maximumsvsvelocities}\caption{Position of the maxima in the outgoing spectra for differents Rotational velocities, (up)Central Distriution, (Down) Homogeneous Distribution.}
\end{figure}

Maximums position provides information about the wave length of the
majority of the outgoing photons after they interact with the gas cloud, as a photon escapes with more scatterings its wave length would be
farthest than the initial which is $1216$ {\AA}. 
As we can see in Figure ~\ref{figure:maximumsvsvelocities} the
position of the maxima does not change with rotational velocity for
the central distribution while for the homogeneous
distriution there is a variation in the maxima position $X_{m}$ for an optical depth $Log_{10}\tau=5$,
this is because at higher rotational velocities photons escape with
less scatterings (fig:ref), for $v=100$\kms the majority of the photons
still escape with scatterings (double-peaked line) but for $v=200$\kms the number
of photons that escape with a few or any scatterings is higher (single-peaked line), this
is one of the main effects of rotation. \\

@@

The width of the line provide information about the range of the wave lengths of the outgoing photons, in the ideal case in which all the photons escape with the same wave length the outgoing spectrum would be the narrower, therefore as the number of scatterings increase the width increases too (reference).	\\

We use the full width at half maximum (FWHM) in order to study the width of the line for all the models that we propose. We found a dependency of the width with the rotation velocity of the gas cloud, this results are summarize in Figure ~\ref{figure:width}. As velocity increase also the equivalent width increase (for both homogeneous and central distribution), it means that some photons escape with fewer scatterings than the static case but at the same time a few photons escape with more scatterings, this can be seen clearly in Figure. (Nscatt)\\

We also found a dependency with the viewing angles, in particular as the angle increase the width also increase Figure ~\ref{fig:withvstheta}, this was also found in Verhemme et al. 2012 for a rotating disk but the pure effect of rotation can not been study as it is in our case. Following the convention of Verhamme et al 2012 we define $\mu=cos(\theta)$, and we can make a polynomial fit for the $FWHM(\mu)$ for our models, this is resumed in Table ~\ref{table:fits}
% explicar como se mide el ancho para algunos casos
% I dont understand this at all, hope Nscatts plot helps 
 %make comparison with Verhamme et al

\begin{figure*}
    \includegraphics[width=0.45\textwidth]{Width7Central.png}
    \includegraphics[width=0.45\textwidth]{Width7Homogeneous.png}
    \includegraphics[width=0.45\textwidth]{Width6.png}
    \includegraphics[width=0.45\textwidth]{Width6HOM.png}
    \includegraphics[width=0.45\textwidth]{Width5.png}
    \includegraphics[width=0.45\textwidth]{Width5HOM.png}
  \label{figure:width}\caption{Width of the lyman-alpha line for all the models. }
\end{figure*}

\begin{figure*}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifODHOM.png}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifOD.png}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifSpeedsHOM.png}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifSpeeds.png}
  \label{figure:widthvstheta}\caption{Width of the lyman-alpha line for all the models. }
\end{figure*}

% fix this table is too big
\begin{table*}
\begin{center}
\begin{tabular}{cc}\hline
Model & Fit\\
Central, $\tau=10^5, V=200$\kms &$FWHM(\mu)=-149877+84509\mu-17864\mu^{2}-1677\mu^{3}-59\mu^{4}$\\
Central, $\tau=10^6, V=200$\kms &$FWHM(\mu)=-32264+9644\mu-1080\mu^{2}+53.8\mu^{3}-\mu^{4}$\\
Central, $\tau=10^7, V=200$\kms & $FWHM(\mu)=-1791+359\mu-27\mu^{2}+0.9\mu^{3}+0.011\mu^{4}$\\
Hom, $\tau=10^5, V=200$\kms & $FWHM(\mu)=-51882.9+12932.5\mu-54.58\mu^{2}-187.22\mu^{3}+11.66\mu^{4}$ \\
Hom, $\tau=10^6, V=200$\kms & $FWHM(\mu)=-7154649+2413478\mu-305268.4\mu^{2}+17158.92\mu^{3}-361.64\mu^{4}$ \\
Hom, $\tau=10^7, V=200$\kms & $FWHM(\mu)=-51074+9912\mu-721\mu^{2}+23.29\mu^{3}-0.28\mu^{4}$ \\
\hline
\end{tabular}
\caption{Fits for EW models} 
\label{table:fits}
\end{center}
\end{table*}

>>>>

\begin{table}
\begin{center}
\begin{tabular}{ccc}\hline
Velocity (\kms) & $V_{\rm max}$&$0,\ 50,\ 100,\ 200,\ 300$\\
Hydrogen Optical Depth & $\tau_{H} $ & $10^{5},\ 10^{6},\ 10^{7}$\\
Dust Optical Depth & $\tau_{A}$ & $0$,$1$\\
Photons Distributions & & Central, Homogeneous\\
\hline
\end{tabular}
\caption{
Values for the varying input parameters in CLARA. Taking into account
all the possible combinations for these models
} 
\label{table:models}
\end{center}
\end{table}

@@

The central result of this paper is summarized in
Fig.~\ref{fig:DifferentVelocities} where we show that rotation has a
considerable effect on the morphology of the emergent Lyman-alpha line
both in the case where the photons are emitted at the sphere's center
and when they are initialized with an homogeneous distribution all
over the gas volume.

The results for these outgoing spectra are integrated over the whole
sphere, meaning that all the escaping photons were taking into account
regardless of the direction of the outgoing photons. Figure
\ref{fig:DifferentObservers} shows how if one gives a weight to each outgoing
photon according to its direction when escaping the gas distribution
it is posible to detect notable differences in the spectrum for
different viewing angles.

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{7tDifSpeedsZ.png}
  \includegraphics[width=0.45\textwidth]{7tHOMDifSpeeds1.png}
\caption{Shape of the lyman alpha line for
    differents velocities. The left (right) panel shows the central
    (homogeneous) photon distribution. All photons were taken into
    account regardless of their outgoing direction of propagation.
  \label{fig:differentvelocities}}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.45\textwidth]{Observers7t.png}
\includegraphics[width=0.45\textwidth]{Observers7tHOM.png}
\caption{Same as Figure \ref{fig:differentvelocities} but this time
  the differen lines show the line morphology for different polar angles
  $\theta$ with respect to the rotation axis. In this case each photon
  has a weight dependent on their outgoing direction. All spectra correspond
  to the same $V_{\rm max}=300$\kms, Optical   Depth $\tau=10^{7}$ and
  Central Distribution without dust.}  
  \label{fig:differentobservers}
\end{figure*}

In the following subsections we quantify the trends observed in
Fig. \ref{fig:differentvelocities} and
Fig. \ref{fig:differentobservers} as a function of the maximum
rotation velocity $V_{\rm max}$ and the position of the observer with
respect to the rotation angle, $\mu = \cos{\theta}$. All the results
in this section will be expressed in terms of frequencies expressed in
velocities.

Next, we  quantify the line by its full width at half maximum
(FWHM) and the peak positions. In the next subsection we take a look
at the escape fraction. We conclude the section by estimating the
expected line flux for top hat filters at a fixed center and varying
width.

\subsection{Line width and peak maxima}
\label{sec:widthpeak}

\begin{figure*}
    \includegraphics[width=0.45\textwidth]{Width7Central.png}
    \includegraphics[width=0.45\textwidth]{Width7Homogeneous.png}
    \includegraphics[width=0.45\textwidth]{Width6.png}
    \includegraphics[width=0.45\textwidth]{Width6HOM.png}
    \includegraphics[width=0.45\textwidth]{Width5.png}
    \includegraphics[width=0.45\textwidth]{Width5HOM.png}
\caption{Width of the lyman-alpha line for all the models.   \label{fig:widthvsvelocity}}
\end{figure*}

\begin{figure*}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifODHOM.png}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifOD.png}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifSpeedsHOM.png}
    \includegraphics[width=0.45\textwidth]{WidthvsThetaDifSpeeds.png}
\caption{Width of the lyman-alpha line for all the models. 
  \label{figure:widthvstheta}
}
\end{figure*}

The first quantitative conclusion of the effect of rotation in the
Lyman alpha line is that the double peaks in the line tend to broaden
until they reach a single broad emission peak. This is most evident in
the case of Lyman-alpha sources homogeneously distributed in the gas
distributions (Fig. \ref{differentvelocities} right panel).

To quantify the line broadening we measure a modified version of the full
width at half maximum (FWHM) for half of the line, $W_{1/2}$. It means that in
the case of double peaked emission, $W_{1/2}$ corresponds to the with
one of thepeaks, while in the extreme case when the line is converted
into a single peak, $W_{1/2}$ corresponds to half of the full width.

This definition allows us to quantify the line width both in the cases
of double and single peak emission. Furthermore it has the advantage
that this line width should have a direct observational correspondence
to the observed line feature once the Inter-Galactic Medium (IGM)
effects are taken into account, which have the central effect of
strongly reducing the intensity of the blue peak of the line.

Figure~\ref{fig:widthvsvelocity} summarizes our findings for $W_{1/2}$
as a function of $V_{\rm max}$. The line width increases with the
rotational velocity of the gas cloud. This increase can be of a factor
of $2-3$ with respect to the width with respect to the static
case. This trend is conserved at all optical depths regardless of the
initial source distribution.

This result includes all the outgoing photons, regardless of the
position of the observer. In Figure~\ref{fig:widthvstheta} we
take into account the different positions of the observer in the
measurement of the half-width $W_{1/2}$. From this we conclude
that observers with a line of sight perpendicular to the axis of
rotation (i.e. edge-on in the case of spiral galaxy) tend to meaure
larger line widths than observers aligned with the rotation axis
(i.e. face-on). The influence of the observer position on the line
width, amounting always less than $15\%$ of a difference with respect
to the result that takes into account all the outgoing photons with
the same weight regardless of the relative observer position.

%Both for central and homogeneous distribution of sources, We found a
%dependency of the width with the rotation velocity of the gas cloud,
%this results are summarize in Figure ~\ref{figure:width}. As velocity
%increase also the equivalent width increase (for both homogeneous and
%central distribution), it means that some photons escape with fewer
%scatterings than the static case but at the same time a few photons escape with% more scatterings, this can be seen clearly in
%Figure. (Nscatt)\\

%We also found a dependency with the viewing angles, in particular as
%the angle increase the width also increase Figure
%~\ref{fig:withvstheta}, this was also found in Verhemme et al. 2012
%for a rotating disk but the pure effect of rotation can not been study
%as it is in our case. Following the convention of Verhamme et al 2012
%we define $\mu=cos(\theta)$, and we can make a polynomial fit for the
%$FWHM(\mu)$ for our models, this is resumed in Table ~\ref{table:fits}

\begin{figure*}
    \includegraphics[width=0.45\textwidth]{maximumvsvelocitiesDifODCentral.png}
    \includegraphics[width=0.45\textwidth]{maximumvsvelocitiesDifODHOM.png}
\caption{Position of the maxima
    in the outgoing spectra for differents Rotational velocities,
    (up)Central Distriution, (Down) Homogeneous Distribution.\label{fig:maximumsvsvelocity}} 
\end{figure*}

The second feature in the line that we use to quantify the effect of
rotation is the position of the line maxima. These provide information
on the wavelength of the majority of the outgoing photons after
they interact with the neutral hydrogen atoms in the gas cloud. If
most of the photon escapes with a low number of scattering, its
outgoing frequency will be close to its initial frequency, that is in
the center of the line. On the contrary if the number of scatterings
is large for the average photon, its outgoing frequency will be far
from the line center. Such reasoning can be made more quantitative to
understand the dependence of the peak maxima as a function of the
hydrogen optical depth in the cloud [citation needed].

In Figure ~\ref{figure:maximumsvsvelocities} we present the position
of the maxima as a function of $V_{\rm max}$. For the photons emitted
in the line center we do not find any variation in the position of the
maxima in the range of explored parameter space. However in the
homogeneous case we can see how the maxima goes to $x_{m}=0$, meaning
that the double peak is converted into a single peak.

This transition to a single peak line occurs for the systems where it
becomes easiear for a bulk of the photons to escape with the lowest
number of scatterings possible. This can explain how the single peak
stage can be achieved in the homogeneous source distribution where
there is fraction of the photons inside a photosphere region with
$\tau_{H}\sim 1$ which allows them to escape within one
scatter. Increasing the rotational velocity makes it easier for the
photons in this photosphere region to escape.

%rotational velocities photons escape with
%less scatterings (fig:ref), for $v=100$\kms the majority of the photons
%still escape with scatterings (double-peaked line) but for $v=200$\kms
%the number of photons that escape with a few or any scatterings is higher
%(single-peaked line), this is one of the main effects of rotation. \\

@@

% explicar como se mide el ancho para algunos casos
% I dont understand this at all, hope Nscatts plot helps 
 %make comparison with Verhamme et al

%\begin{figure}
%    \includegraphics[width=0.45\textwidth]{maximumvsODDifSpeedsCentral.png}
%    \includegraphics[width=0.45\textwidth]{maximumvsODDifSpeedsHOM.png}
%  \label{fig:maximavsOD}\caption{Position of the maxima in the
%    outgoing spectra for differents Optical Depths, (up)Central
%    Distriution, (Down) Homogeneous Distribution.} 
%\end{figure}

% fix this table is too big
%\begin{table*}
%\begin{center}
%\begin{tabular}{cc}\hline
%Model & Fit\\
%Central, $\tau=10^5, V=200$\kms &$FWHM(\mu)=-149877+84509\mu-17864\mu^{2}-1677\mu^{3}-59\mu^{4}$\\
%Central, $\tau=10^6, V=200$\kms &$FWHM(\mu)=-32264+9644\mu-1080\mu^{2}+53.8\mu^{3}-\mu^{4}$\\
%Central, $\tau=10^7, V=200$\kms & $FWHM(\mu)=-1791+359\mu-27\mu^{2}+0.9\mu^{3}+0.011\mu^{4}$\\
%Hom, $\tau=10^5, V=200$\kms & $FWHM(\mu)=-51882.9+12932.5\mu-54.58\mu^{2}-187.22\mu^{3}+11.66\mu^{4}$ \\
%Hom, $\tau=10^6, V=200$\kms & $FWHM(\mu)=-7154649+2413478\mu-305268.4\mu^{2}+17158.92\mu^{3}-361.64\mu^{4}$ \\
%Hom, $\tau=10^7, V=200$\kms & $FWHM(\mu)=-51074+9912\mu-721\mu^{2}+23.29\mu^{3}-0.28\mu^{4}$ \\
%\hline
%\end{tabular}
%\caption{Fits for EW models} 
%\label{table:fits}
%\end{center}
%\end{table*}
